\documentclass{article}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{cite}
\begin{document}
\lstset{language=python, tabsize=4}
\title{CS 516 homework 3.0}
\author{Yuchen Hou}
\maketitle

\section{A-TM-abba is undecidable}
Proof by reduction(from A-TM to A-TM-abba) and contradiction(assumption: A-TM-abba is decidable). From the assumption, we can construct decider N(input = [M], where M is a TM) to decide A-TM-abba:
\begin{enumerate}
	\item if M accepts abba: accept [M]
	\item if M doesn't accept abba: reject [M]
\end{enumerate}
We can construct decider O(input = [P, w], where P is a TM and w is a word) to decide A-TM:
\begin{enumerate}
	\item construct TM Q(input = x, where x is a word) to accept abba when M accepts w:
	\begin{enumerate}
		\item run M on w
		\item if M accepts w: accept x
	\end{enumerate}
	\item run N on [Q]
	\item if N accepts [Q]: accept [P, w]
	\item if N rejects [Q]: reject [P, w]
\end{enumerate}
This is a contradiction with A-TM is undecidable. Therefore, A-TM-abba is undecidable.

\section{E-jump-LBA is undecidable}
Proof by reduction(from A-2CM to E-jump-LBA) and contradiction(assumption: E-jump-LBA is decidable). From the assumption, we can construct decider N(input = [M], where M is a jump-LBA) to decide E-jump-LBA:
\begin{enumerate}
	\item if L(M) is empty: accept [M]
	\item if L(M) is not empty: reject [M]
\end{enumerate}
We can construct decider O(input=[P], where P is a 2CM) to decide A-2CM:
\begin{enumerate}
	\item construct jump-LBA Q(input = [w], where w is a word) to accept accepting computation histories of P:
	\begin{enumerate}
		\item assert w has valid computation history format: [c[1], c[2], ..., c[l]], where c[i] = [instruction, state, counter1, counter2]
		\item assert c[1] is the initial configuration for P
		\item assert all configuration transitions are valid (mark, jump and move to next symbol to check adjacent configurations: read instruction, check state and counter updates)
		\item assert c[l] is an accepting configuration for P
		\item accept w
	\end{enumerate}
	\item Run N on [Q]
	\item if N rejects [Q]: accept [P]
	\item if N accepts [Q]: reject [P]
\end{enumerate}
This is a contradiction with A-2CM is undecidable. Therefore, E-jump-LBA is undecidable.

\section{Language is in RE}
As L is in R, we can construct decider M(input = [w], where w is a word) to decide L. Therefore we can construct TM N(input = x) to accept L':
\begin{enumerate}
	\item guess word y
	\item run M on yxy
	\item if M accepts yxy: accept x
\end{enumerate}
As TM N accepts L', L' is in RE.

\section{Reverse of R is R}
As L is decidable, we can construct decider M to decide L. Therefore we can construct decider N(input = x) to decide $ L^r $:
\begin{enumerate}
	\item reverse x in place to get $ x^r $ by swapping corresponding symbol pairs
	\item run M on $ x^r $
	\item if M accepts $ x^r $: accept x
	\item if M rejects $ x^r $: reject x
\end{enumerate}
As decider N decide $ L^r $, $ L^r $ is decidable.

\section{E-read-only-2-TM is undecidable}
Proof by reduction(from A-2CM to E-read-only-2-TM) and contradiction(assumption that E-read-only-2-TM is decidable). From the assumption, we can construct decider N(input = [M], where M is read-only-2-TM) to decide E-read-only-2-TM:
\begin{enumerate}
	\item if L(M) is empty: accept [M]
	\item if L(M) is not empty: reject [M]
\end{enumerate}
We can construct decider O(input = [P], where P is a 2CM) to decide A-2CM:
\begin{enumerate}
	\item construct read-only-2-TM Q(input = [w], where w is a word) to accept anything when P halts:
	\begin{enumerate}
		\item simulate P with its head coordinates simulating the counters
		\item if the simulation finishes, accept w
	\end{enumerate}
	\item run N on [P]
	\item if N rejects [Q]: accept [P]
	\item if N accepts [Q]: reject [P]
\end{enumerate}
This is a contradiction with A-2CM is undecidable. Therefore, E-read-only-2-TM is undecidable.

\end{document}
